### 2025-05-23 docs/manuscript/manuscript_outline.Rmd
add manuscript_outline

---
title: "manuscript_outline"
output: html_document
date: "2025-05-23"
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
library(hatchR); library(tidyverse); library(patchwork)
```

# Title Page

A generalizable tool for predicting developmental phenology for wild poikilotherms.

Morgan M. Sparks, Brian Leavell, Bryan M. Maitland

# Abstract

Before exogenous feeding, poikilothermic organisms have a near mechanistic relationship between ambient temperature and developmental rates. As such, statistical models can be easily developed to predict when organisms develop. Until recently, most models only used non-variable developmental regimes making them difficult to apply to wild environments. However, the R package hatchR formalized an approach using effective values where each day is given a developmental unit, accurately predicting developmental phenology for wild poikilotherms. hatchR was developed specific to fish, however this manuscripts broadens the tool's application showing how it can be used to predict developmental phenology for a broad range of taxa, including amphibians, reptiles, and invertebrates. Moreover, we provide numerous examples of how this approach informs scenarios from applied management to basic questions regarding ecological and evolutionary questions.

# Introduction

P1: Poikilotherms and ambient temperatures

P2: ATU models and effective value models

P3: Examples with other species

P4: Our approach here (R package and Shiny app)

# Methods

## Effective value models

Effective value models function by leveraging the statistical relationship derived from raising poikilotherms at different temperatures and fitting that relationsip with a non-linear model. The formultation of that relationship can then be reciprocated which provides the unit of development for a day's average temperature--an effective value. Effective value models function then by cumulatively summing to one at which the organism achieves the development of the parameterized trait.

The model follows the general format of:

$$
Effective Value_i = 1/exp(log_ea - log_e(Temperature_i - b))
$$

Where *i* is the daily value and a fish hatches or emerges when the cumulative sum reaches one: $$\sum_{i =1}^nEffectiveValue_i = 1$$

As an example, we parameterize an effective value model for western tailed frogs (*Ascaphus truei*) common to western North America (Figure 1). Custom parameterized models use the `fit_model()` function in hatchR, which is built on model 2 using the power law from @beacham1990. Alternatively, to actually predict phenology using hatchR, the `predict_phenology()` only expects a model expression as input and could assume other model formulations custom built outside of the package by the user.

```{r}
# Brown 1975, Comparitive Biochemistry and Physicology Part A: Physiology
# https://www.sciencedirect.com/science/article/pii/030096297590033X

### parameterize mod
ascaphus_data <- tibble(temp = c(7.6,9.8,11,13,14.5,15,18),
                        days = c(44,27.1,22.6,16.1,13.4,12.7,10.7))

ascaphus_mod <- fit_model(temp = ascaphus_data$temp,
                          days = ascaphus_data$days,
                          species = "ascaphus",
                          development_type = "hatch")

### get effective values
temps <- c(6:20)

ef_vals <- NULL
for (x in temps) {
  ef <- eval(parse(text = ascaphus_mod$expression$expression))
  ef_vals <- rbind(ef_vals, ef)
}

ascaphus_ef <- matrix(NA, 15, 2) |> tibble::tibble()
colnames(ascaphus_ef) <- c("temp", "ef")
ascaphus_ef$temp <- temps
ascaphus_ef$ef <- ef_vals[, 1]

### plot

p1 <- ascaphus_mod$pred_plot +
  labs(x = "Incubation Temperature (°C)", y = "Days to Hatch") +
  lims(y = c(0, 50))

data_1 <- tibble(t = c(0:20), e = seq(0, 0.20, by = 0.01))

p2 <- ascaphus_ef |>
  ggplot() +
  geom_point(aes(x = temp, y = ef)) +
  geom_line(aes(x = temp, y = ef)) +
  geom_line(data = data_1, aes(x = t, y = e), linetype = "dashed") +
  # geom_abline(intercept = 0, slope = .01, linetype = "dashed") +
  labs(x = "Daily average temperature (°C)", y = "Effective value") +
  theme_classic()

p1 + p2
```

## Data and data checks

hatchR requires two essential paired vectors of data, one of daily average temperature and the other the date for those temperatures. The software is designed to function around common field temperature loggers and provides users the ability to summarize temperatures with multiple daily recordings.

### Data input

Maybe include the below:

```{r}
df <- data.frame(
  date = c(
    "2000-01-01",
    "...",
    "2000-07-01",
    "...",
    "2000-12-31"
  ),
  temperature = c(
    2.51,
    "...",
    16.32,
    "...",
    3.13
  )
)

knitr::kable(df,
  booktabs = TRUE,
  caption = "Example temperature data for use in hatchR."
)
```

If you import data from raw files with multiple daily readings, the package allows you to summarize your with `summarize_temp()` and then check summarized data with the `plot_check_temp()` and `check_continuou()` functions.

# Case studies

There are numerous applications for effective value models and poikilotherms that expand well beyond the bounds of this manuscript. We provide XXX examples in the table below of studies that could be used to parameterize custom models using the `fit_model()` function from hatchR. These are a non-exhaustive search of the both peer-reviewed and grey literature, but demonstrate the wide taxonomic spread that could be paired with the effective value approach.

## Ecological

Tailed frog and westlope cutthroat example???

## Phylogenetic

```{r}

### all the Rana's below refer to taxonomy from the paper, they are now genus Lithobates

# Rana sylvatica ----------------------------------------------------------

# Moore, 1939 Ecology
# https://www.jstor.org/stable/1930439?seq=1

rana_s_data <- tibble(temp = c(9, 21 ,26),
                      days = c(15,13,4))

rana_s_data <- tibble(temp = c(10,15.3,18.5,19.9, 23.7),
                      hours = c(305,128,87,79,52)) |>
  mutate(days = ceiling(hours/24))

rana_s_mod <- fit_model(temp = rana_s_data$temp,
                        days = rana_s_data$days,
                        species = "Rana_sylvatica",
                        development_type = "hatch")

# Rana pipiens -----------------------------------------------------------
# from Moore 1939

rana_pip_data <- tibble(temp = c(15.3,18.6,19.8, 26),
                        hours = c(168,116,99,57)) |>
  mutate(days = ceiling(hours/24))

rana_pip_mod <- fit_model(temp = rana_pip_data$temp,
                          days = rana_pip_data$days,
                          species = "Rana_pipiens",
                          development_type = "hatch")

# Lithobates clamitans ----------------------------------------------------------
# from Moore 1939

rana_c_data <- tibble(temp = c(15,19.8, 25.3,33.4),
                      hours = c(263,112,62,45)) |>
  mutate(days = ceiling(hours/24))

rana_c_mod <- fit_model(temp = rana_c_data$temp,
                        days = rana_c_data$days,
                        species = "Lithobates_clamitans",
                        development_type = "hatch")

# Rana palustris ----------------------------------------------------------
# from Moore 1939

rana_p_data <- tibble(temp = c(15.5,18.6, 19.9, 25.7),
                      hours = c(190,126,106,63)) |>
  mutate(days = ceiling(hours/24))

rana_p_mod <- fit_model(temp = rana_p_data$temp,
                        days = rana_p_data$days,
                        species = "Rana_palustris",
                        development_type = "hatch")

# Lithobates example --------------------------------------------------------------------

set.seed(123)
temperature_data <- data.frame(temperatures = sort(rnorm(30, mean = 16, sd =2.5)),
                               dates = seq(ymd("2000-04-01"), ymd("2000-04-30"), by = "days"))

froggy_mods <- c(rana_s_mod$expression$expression,
                 rana_p_mod$expression$expression,
                 rana_c_mod$expression$expression,
                 rana_pip_mod$expression$expression)

froggy_hatch <- map(froggy_mods,
                    predict_phenology,
                    data = temperature_data,
                    dates = dates,
                    temperature = temperatures,
                    spawn.date = "2000-04-01") |>
  map_dbl("days_to_develop") |>
  sort(decreasing = TRUE) |>
  tibble()

colnames(froggy_hatch)[1] <- "days_to_develop"

froggy_hatch <- froggy_hatch |>
  mutate(start = ymd("2000-04-01"),
         stop = start + days_to_develop,
         species = c("Green Frog", "Pickerel Frog", "Northern Leopard Frog", "Wood Frog"),
         index = seq(20.5,17.5, by = -1))

ggplot() +
  geom_point(data = temperature_data, aes(x = dates, y = temperatures)) +
  geom_line(data = temperature_data, aes(x = dates, y = temperatures)) +
  geom_rect(data = froggy_hatch, aes(xmin = start, xmax = stop, ymax =index-.15, ymin = index-.5, fill = species)) +
  geom_label(data = froggy_hatch, aes(x = start + (stop - start) / 1.45, y = (index -0.325), label = days_to_develop)) +
  labs(x = "Date", y = "Temperature (°C)",
       title = "Development periods of four North American Frog Species") +
  scale_fill_manual(values = c("darkolivegreen4","deepskyblue4", "grey23",  "purple4"), name = "Species") +
  theme_classic() +
  theme(legend.position = c(0.75, 0.25))

```

## Local Adaptation

Need an example where multiple populations from the same species are parameterized. ideally not an amphibian

# Discussion

P1: Summary of above

P2: Caveats

-   namely that some taxa like insects and frogs are much more likely to bail on development when environmental cues suggest they need to

P3: Other considerations?

P4: ???

### Conclusion

# Bibliography